---
title: "Base layer proof of concept"
author: "Maximiliano Micheli"
date: "15 March 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=TRUE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(sp)
library(raster)
library(rgeos)

library(viridis)
library(gridExtra)
library(mapview)
library(stars)
library(sf)
set.seed(1)

```

## Why a basic spatial unit (or base layer)?

When performing an intersection of two layers, there are certain part of the polygons that are lost. Imagine the following two polygons, which are close to be the same but they are in slightly different places

```{r}
set.seed(131)
library(sf)
m = rbind(c(0,0), c(1,0), c(1,1), c(0,1), c(0,0))
p = st_polygon(list(m))
n = 2
l = vector("list", n)
for (i in 1:n)
  l[[i]] = p + 0.25 * runif(2)
s = st_sfc(l)
plot(s, col = sf.colors(categorical = TRUE, alpha = .5))
title("overlapping squares")
```

When doing an intersection (red colour) of these two, the resulting area is smaller than the original area of both squares. This mean that part of this area is lost in the process. If we add another slightly different layer, probably the total area will change. Moreover, the shape of the result will change, which will lead to different shapes depending on what we are intersecting. 

```{r}
sf_s <- st_as_sf(s)
inter <- st_intersection(sf_s[1,],sf_s[,])
sf_s2<- rbind(sf_s,inter)
plot(sf_s2, col = sf.colors(categorical = TRUE, alpha = .5))
```

If there was a Basic Spatial Unit (BSU), that confidently represents the area we want to work with, we wouldn't care of areas that lay outside the BSU. Having a graph of this, we can see that our two shapes now overlay in a BSU, where all the features get captured in each cell. This allows that any attributes that we add in the future are consistent spatially to allow comparisons over time. The BSU always stays the same, and each of the cells in this grid absorbs the attributes of the overlaying shapes.


```{r}
m = rbind(c(0,0), c(1.25,0), c(1.25,1.25), c(0,1.25), c(0,0))
p = st_polygon(list(m))
s = st_sfc(p)
grid <- st_make_grid(s)

vec <- c(1,2)
sf_s$cat <- vec
mapview(grid)+mapview(sf_s, zcol ='cat')
```


## Using SA2 as a base layer

Here we upload our SA2 layer, transform it to a meters CRS and filter a few SA2 codes to keep work simple

```{r , message = FALSE, echo = FALSE, results = FALSE}
# Uploading SA2
SA2 <- st_read('C:/Users/Maxi/Documents/GitHub/Base-Layer/Datasets/SA2_2021_GDA94/SA2_2021_AUST_GDA94.shp')

# Uploading ARKS
ARKS <- st_read('C:/Users/Maxi/Documents/GitHub/Base-Layer/Datasets/Biodiversity_KoalaPrioritisationProjectNSW_ARKS/KoalaPrioritisationProjectNSW_ARKS.shp')
ARKS <- st_transform(ARKS, crs = 3577)

# Transform CRS
SA2$AREASQKM21[SA2$STE_CODE21 == 1]
SA2 <- st_transform(SA2, 3577)
SA2$area_calc <- st_area(SA2)

# Filter NSW
SA2 <- SA2[SA2$STE_CODE21 == 1,]

# Testing for some SA2
port <- SA2[SA2$SA2_CODE21 == '106031119'|SA2$SA2_CODE21 == '106031120'|SA2$SA2_CODE21 == '106031121'|SA2$SA2_CODE21 == '106031124'|SA2$SA2_CODE21 == '106031125'| SA2$SA2_CODE21 == '106031132'|SA2$SA2_CODE21 == '106031122' | SA2$SA2_CODE21 =='111031232',]
#
```



## Plot of selected area

Map of the selected areas for base layer 

```{r pressure, echo=FALSE}
mapview(port)
```


## Making a grid of the selected area

This poses some questions that is good to answer at this stage of what we are doing:

- Is the selection of SA2 as the base layer a good choice?
Criteria:
1) Stability in time and space - Our base layer should serve to the purpose of bringing stability among all of our products and to be stable in time. 
2) Accuracy in the description state's territory: the base layer should be reliable, since all the area calculations and comparisons are going to be based on this territory. 

- Is the size of the grid relevant? Should we build a small grid (with high computing costs but very detailed information) or a big grid (with low computing costs but not too detailed information). For example, if we have a grid with 1km cells and land use with 50m cells, this will lead to high imprecision when aggregating land use to our 1km cells. 

- Is the form of the grid relevant? 
We need to analyze if there is a trade off between using squared grid or hexagonal grid

```{r}

# test
port <- st_cast(port, "POLYGON")
port$area_calc <- st_area(port)
BaseGrid <- st_make_grid(port, 250, square = FALSE)
mapview(BaseGrid)

```

Now the displayed grid contains all the SA2 areas. The grid cells are partitioned at some point. They are not comparable between them. However, if the structure stays the same they are comparable in time (if al use this same BSU) and with other layers that use the same BSU. 

On the other hand, having cells that are not complete generates a problem of aggregation (unless we have an attribute such as SA2).

```{r}

#test$area_calc = st_area(test)
#mapview(test)
#sum(test$area_calc)
#sum(port$area_calc)

new <- st_intersection(st_as_sf(BaseGrid),port)
new$area_calc <- st_area(new)
mapview(new)


```

```{r, include = FALSE}

sum(port$area_calc)
sum(new$area_calc)
```

```{r, include = FALSE}
############ testeando los resultados #################
ver1 <- port %>% 
  st_drop_geometry %>% 
  dplyr::group_by(SA2_CODE21) %>% 
  dplyr::summarize(area_original = sum(area_calc))

ver2 <- new %>% 
  st_drop_geometry %>% 
  dplyr::group_by(SA2_CODE21) %>% 
  dplyr::summarize(area_hexagon = sum(area_calc))

comparison <-merge(ver1,ver2, by = 'SA2_CODE21') ### port 2 is a solid candidate for grid
comparison$ratio <- comparison$area_original/comparison$area_hexagon

```

We can see that our new hexagonal layer has the same overall area for all the SA2 and for each of them in the original file. This means it keeps the area constant, which is a good outcome in terms of precision. 

The trade-off in this case, is that our hexagons are partitioned at some point, so we won't keep them in it's original shape. In terms of our product, it doesn't change much. We've now got a base layer that accurately represents our original file.

```{r}
comparison
```

# Adding another layer to our BaseLayer

```{r}
new_layer <- st_join(new,ARKS[,c('PoRS_Name')], left = T, largest = TRUE )
new_layer$geometry_type <- st_geometry_type(new_layer)
new_layer <- new_layer[new_layer$geometry_type!='POINT',]
#mapview(ARKS, col.regions = 'red')
```

```{r}
table(new_layer$geometry_type)
```


Checking that areas remain the same
```{r}
sum(new_layer$area_calc)
sum(st_area(new_layer))
```

Checking that internal areas for SA2 stay the same

```{r, include = FALSE}
############ testeando los resultados #################
ver1 <- port %>% 
  st_drop_geometry %>% 
  dplyr::group_by(SA2_CODE21) %>% 
  dplyr::summarize(area_original = sum(area_calc))

ver2 <- new_layer %>% 
  st_drop_geometry %>% 
  dplyr::group_by(SA2_CODE21) %>% 
  dplyr::summarize(area_hexagon = sum(area_calc))

comparison <-merge(ver1,ver2, by = 'SA2_CODE21') ### port 2 is a solid candidate for grid
comparison$ratio <- comparison$area_original/comparison$area_hexagon

```

```{r}
comparison
```

Checking results for ARKS Areas

```{r}
mapview(new_layer)+ mapview(ARKS, col.regions = 'red')


```

We can see that the share of ARKS will change after intersecting with the new layer, because the base layer doesn't contain areas that the ARKS area has. So this is a risk

- If an area of an attribute layer falls outside of our base layer, what do we do to that?
- As we are working with an intersection method, a cell will get the attributes of another layer if a very tiny portion of the latter intersects the former. 
- When we intersect with more than three features on a cell, we are keeping the feature that intersects in it's largest area

